### 2018-04-09 paper/abstract.tex

\begin{abstract}
  % problem, context
  Recursion-free Constrained Horn Clauses (CHCs) are logic-programming
  problems that formulate verifying safety of programs with bounded
  iteration and recursion.
  %
  They both formulate bounded verification problems and are generated
  by solvers that attempt to verify safety of unbounded iterative or
  recursive systems.
  %
  Efficient solvers of recursion-free systems reduce the problem to
  solving a series of systems that can each be solved tractably.

  % contribution:
  In this paper, we define a novel tractable class of recursion-free
  systems, named \emph{Clause-Dependence Disjoint (CDD)} systems, that
  strictly generalizes tractable classes defined in previous work.
  %  
  We implemented a novel solver for recursion-free CHCs, named \sys,
  that reduces the problem of solving a given recursion-free system to
  solving an equivalent CDD system.
  %
  Our evaluation of \sys on standard benchmarks indicates that, in
  many practical cases, it performs significantly better than a
  state-of-the-art implementation.
\end{abstract}