### 2018-04-09 paper/fig/bench.tex

\begin{center} 
  \begin{table*} 
    \centering
    \begin{tabular}{| r| r | r | r | r |} 
      \hline
      \multicolumn{1}{|c }{ \textbf{Benchmarks}}
       & \multicolumn{2}{|c|}{\sys} & 
       \multicolumn{2}{c|}{\duality} \\
       \hline
       \multicolumn{1}{|c|}{\textbf{Number}} 
       &  \multicolumn{1}{|c|}{\textbf{2 X}} 
       &  \multicolumn{1}{c|}{\textbf{TO}} 
       &  \multicolumn{1}{|c|}{\textbf{2 X}} 
       &  \multicolumn{1}{c|}{\textbf{TO}} 
       \\
      \hline
       670 & 194 & 19 & 24 & 8\\
      \hline
    \end{tabular}
    % 
    \caption{The results of our evaluation of \sys.
      % 
      "Benchmarks'' contains features of the
      benchmarks, in particular the 
       number of CHC problems ("Number")
      % 
      "\sys" contains features of the
      performance of \sys, in particular  ("2 X") contains the number of 
      CHC problems that at least two times faster than \duality , ("TO") contains the 
      number of CHC problems that time out.
      %
       " \duality " contains features of the
      performance of a, in particular  ("2 X") contains the number of 
      CHC problems that at least two times faster than \sys, ("TO") contains the 
      number of CHC problems that time out.
      %
      When we compare the running time, we eliminate all cases if both solver can solve it under 0.1 second
      }
    \label{table:bench}
  \end{table*} 
\end{center}