### 2018-04-10 paper/fig/hierarchy.tex

\scalebox{0.75}{%
\begin{tikzpicture}
\begin{scope}
  \node (top) at (0,0) {\begin{tabular}{c}All Systems\\ \textsc{Undecidable}\end{tabular}};
    \node[below = 0.5 of top] (rec) {\begin{tabular}{c}Recursion Free\\
  \textsc{co-Nexptime Complete}\end{tabular}};
  \node[below = 0.5 of rec] (dir) {\begin{tabular}{c}Directly Solvable\\ \textsc{co-NP}\end{tabular}};
  \node[label={0:\{\textsc{Shara}\}}, draw, rounded corners=0.5cm,
      below = 0.5 of dir] (cdd) {\begin{tabular}{c}Clause-Dependent-Disjoint\\ \textsc{co-NP}\end{tabular}};
  \node[label={0:\{\textsc{Ufo}\}},
      below left = 0.5 and 0.5 of cdd] (lin) {\begin{tabular}{c}Linear\\ \textsc{co-NP}\end{tabular}};
  \node[label={180:\{\textsc{Eldarica}\cite{rummer13b}\}},
      below right = 0.5 and 0.5 of cdd] (bd) {\begin{tabular}{c}Body Disjoint\\ \textsc{co-NP}\end{tabular}};
  \node[label={180:\{\textsc{Duality}\cite{bjorner13}\}},
      below = 0.5 of bd] (tree) {\begin{tabular}{c}Tree\\ \textsc{co-NP}\end{tabular}};
\end{scope}
\begin{scope}[every edge/.style={draw=black, very thick}]
  \path[-] (tree) edge (bd);
  \path[-] (bd) edge (cdd);
  \path[-] (lin) edge (cdd);
  \path[-] (cdd) edge (dir);
  \path[-] (dir) edge (rec);
  \path[-] (rec) edge (top);
\end{scope}
\end{tikzpicture}
}