### 2018-04-11 paper/abstract.tex
pass on first half of paper

\begin{abstract}
  % problem, context
  Recursion-free Constrained Horn Clauses (CHCs) are logic-programming
  problems that can model safety properties of programs with bounded
  iteration and recursion.
  %
  In addition, many recursive CHC system solvers reduce inputs to a
  series of recursion-free CHC systems that can each be solved
  efficiently.

  % contribution:
  In this paper, we define a novel class of recursion-free systems,
  named \emph{Clause-Dependence Disjoint} (CDD), that
  generalizes classes defined in previous work.
  %
  The advantage of this class is that many systems are significantly
  smaller than equivalent systems in other classes which allows them
  to be solved more efficiently.
  %
  Based on the class of CDD systems, we implemented a CHC solver named
  \sys.  \sys~solves arbitrary CHC systems by reducing the input to a
  series of CDD systems.
  %
  Our evaluation indicates that \sys~outperforms state-of-the-art
  implementations in many practical cases.
\end{abstract}