### 2018-04-11 paper/overview.tex

\section{Overview}
\label{sec:overview}

%first introduce the question
In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
$\mcchc$, that models the safety of a small program,
\cc{dblAbs} (\autoref{fig:multicall-code}).
%
In \autoref{sec:solve-ex}, we show that $\mcchc$ is a CDD system and
how \sys can solve it by encoding it into binary interpolants.
%
In \autoref{sec:not-in}, we illustrate that $\mcchc$ is not in
directly solvable classes introduced in previous work.
%

% introduce running example program
\subsection{Verifying \cc{dblAbs}: an example hierarchical program}
\label{sec:running-ex}

% include code of running example and some figure:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.26\textwidth] %
    { \caption{\cc{dblAbs}: an example hierarchical program.} %
      \label{fig:multicall-code} }
    { \input{code/dblAbs.java} }
    %
    \ffigbox[.7\textwidth] %
    {\caption{$\mcchc$ as a directed hypergraph.
        % 
        Each relational predicate is depicted as a graph node
        while each clause is represented by a hyperedge.
        Each hyperedge is labelled by the fact in the corresponding CHC\@.
        Each node has a valid corresponding interpretation,
        written in braces.
      } %
      \label{fig:ex-graph} }
    % { \includegraphics[width=\linewidth]{fig/ex-deps.pdf} }
      {\input{fig/example}}
  \end{floatrow}
\end{figure}
% walk through code of running example:
%
\cc{dblAbs} is a small procedure that doubles the absolute value of
its input and stores the result in \cc{res}.
% state the safety problem:
The program also asserts that \cc{res} is greater than or equal to $0$
before exiting.
% talk about solving Horn Clauses:
Verifying this assertion reduces to solving a recursion-free
CHC system over a set of uninterpreted predicates that represent
the control locations in \cc{dblAbs}.
%
In particular, one such system $\mcchc$ is
%
\begin{align}
\label{chc}
  % semantic constraint of dbl:
\cc{dbl}(\cc{x}, \cc{d}) &\gets \cc{d} = 2 * \cc{x} \\
  % semantic constraint of line 3:
  \cc{L}_4(\cc{n}, \cc{abs}) &\gets  \cc{abs} = 0 \\
  % semantic constraint of line 5 (then branch)
  \cc{L}_6(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} \ge 0 \\
  % semantic constraint of line 6 (else branch)
  \cc{L}_8(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} < 0 \\
  % assign branch
  \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_6(\cc{n},\cc{abs}) \land \cc{abs'} = \cc{n} \\
   \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_8(\cc{n},\cc{abs}) \land \cc{abs'} = -\cc{n} \\
  % semantic constraint of main procedure:
   \cc{main}(\cc{n},\cc{res}) &\gets \cc{L}_9(\cc{n}, \cc{abs'}) \land
                                     \cc{dbl}(\cc{x}, \cc{d})
    \land \cc{abs'} = \cc{x} \land \cc{res} = \cc{d} \\
  %
    \bot &\gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
\end{align}
%
While $\mcchc$ has been presented as the result
of a translation from \cc{dblAbs}, \sys is
purely a solver for CHC systems: it does not require access to the
concrete representation of a program, or for a given CHC system to be
the result of translation from a program at all.

% solving the running example
\subsection{$\mcchc$ as a Clause-Dependence Disjoint System}
\label{sec:solve-ex}
% state it is CDD
The recursion-free CHC system $\mcchc$ is a
\emph{Clause-Dependence-Disjoint} (CDD) system.
% state the defn of CDD:
A CHC system can be classified as CDD when the clauses obey
rules based on dependencies between the predicates.
%
One predicate is directly dependent on another if the second appears
in the body of a clause where the first is the head.
%
Based on this, we say that a clause is dependence disjoint if
\textbf{(1)} no two predicates in the body share any transitive
dependencies on other predicates and \textbf{(2)} no two predicates in
the body are the same.
% give a example
As an example, clause $(7)$ is dependence disjoint. Clause $(7)$ uses
both $\cc{L}_9$ and $\cc{dbl}$ in its body.
%
The transitive dependency of $\cc{L}_9$ is the set
$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\}$ while the transitive dependency of
\cc{dbl} is the empty set.
%
Therefore, their transitive dependencies are disjoint:
$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\} \cap \varnothing = \varnothing$.
%
All other clauses in $\mcchc$ have at most one uninterpreted predicate
in the body, so they are trivially disjoint dependent.
%
When each clause in a CHC system is dependence disjoint, it is a CDD
system. Therefore $\mcchc$ is a CDD system.
%
%
The formal definition of CDD and its key properties are given in
\autoref{sec:CDD-defn}.
The formal definition of transitive dependency is given in\QZ{reference}.
%

% restate defn of CDD systems:
\sys solves CDD systems directly by issuing a binary interpolation
query for each uninterpreted predicate in topological order.
%
Each uninterpreted predicate $P$ has a valid interpretation which can
be computed by interpolating
\textbf{(1)} the \emph{pre}-formula, constructed from clauses where
$P$ is the head and
\textbf{(2)} the \emph{post}-formula, constructed from
all clauses where the head transitively depends on $P$.
%

We will walk through the construction of an interpolation query for
$\cc{L}_9$ by hand.  By the time \sys reaches $\cc{L}_9$ it will
already have solutions for $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$.
Possible interpretations of these predicates are shown in
\autoref{fig:ex-graph}.
%
A solution for $\cc{L}_9$ can be constructed by interpolating
the appropriate pre and post formulas.
%
The pre-formula is constructed from the bodies of clauses where
$\cc{L}_9$ is the head. Each relational
predicate, $P$,
is replaced by a corresponding boolean indicator variable, $\cc{b}_P$, and each
boolean indicator variable implies the solution for its predicate.
%
In particular, the pre-formula for $\cc{L}_9$ is constructed from
clauses (5) and (6):
\begin{gather}
  ((\cc{b}_{\cc{L}_6} \land \cc{abs'} = \cc{n})
  \lor
  (\cc{b}_{\cc{L}_8} \land \cc{abs'} = -\cc{n}))
  \land
  (\cc{b}_{\cc{L}_6} \Rightarrow \cc{n} \ge 0)
  \land
  (\cc{b}_{\cc{L}_8} \Rightarrow \cc{n} < 0)
\end{gather}
%
The post-formula is constructed from clauses that transitively depend
on $\cc{L}_9$. Again, we replace relational predicates by
corresponding boolean indicators. However, we omit the boolean indicator for
$\cc{L}_9$. The post-formula is composed from clauses (1), (7), and
(8):
\begin{gather}
  (\cc{b}_{\cc{main}} \land \cc{res'} < 0)
  \land
  (\cc{b}_{\cc{main}} \Rightarrow
    (\cc{b}_{\cc{dbl}}
    \land \cc{abs'}= \cc{x}
    \land \cc{res}=\cc{d} ))
  \land (\cc{b}_{\cc{dbl}} \Rightarrow \cc{d} = 2*\cc{x})
\end{gather}
%
Interpolating the pre and post formulas yields a valid intepretation
of $\cc{L}_9$: $\cc{abs'} \geq 0$.
The procedure for solving a CDD system is described in formal detail
in \autoref{sec:solve-cdd}.

\subsection{$\mcchc$ is not in other recursion-free classes}
\label{sec:not-in}
%
In this section, we show that $\mcchc$ cannot be placed in other
known classes of recursion-free CHC systems. Specifically, we will
discuss body-disjoint systems and linear systems.
%

% tree solver
Tree structure
systems~\cite{mcmillan14,bjorner13,heizmann10,rummer13a,rummer13b} are
class of recursion-free CHC system where each uninterpreted predicate
appears in the body of at most one clause (and appears in such a
clause exactly once).
%
Such systems cannot model a program with multiple control paths that
share a common subpath, typically modeled as a CHC system with a
uninterpreted predicate that occurs in the body of multiple clauses.
%
$\mcchc$ is not a body-disjoint system because the uninterpreted
predicates $\cc{L}_4$ shows in the body of both clause (3) and clause
(4).
%
In order to find an interpretation for $\mcchc$, a solver which uses
tree-structure systems would have to create a fresh copy of
$\cc{L}_4$, as well copies for all relational predicates that
$\cc{L}_4$ depends on.

%linear systems:
Previous work also introduced linear system~\cite{albarghouthi12a},
where each clause's body has at most one uninterpreted predicate.
%
However, such systems cannot directly model the control flow of a
program that contains procedure calls.
%
$\mcchc$ is not a linear system because the body of clause (7) has two
predicates \cc{L9} and \cc{dbl}.
%
CHC solvers that solve linear systems effectively inline the
constraints for relational predicates that occur in non-linear
clauses~\cite{albarghouthi12b}.
%
In the case of $\mcchc$, inlining the constraints of $\cc{dbl}$ is
efficient, but in general such approaches can generate systems that
are exponentially larger than the input.

%% talk about expansion:
\begin{figure}
  \input{fig/hierarchy}
\end{figure}

The class CDD of CDD systems is a superclass of the classes of
body-disjoint and linear systems.
%
% The fact that $\mcchc$ is CDD corresponds to the fact that there no
% statements executed more than once within one execution path.
% \DAH{elaborate}
%
E.g., a recursion-free system that models \cc{dblAbs} 
call \cc{dbl} twice on \cc{abs} in sequence would not
be CDD.
%
\sys, given an arbitrary recursion-free CHC system $S$,
solves it by generating an equivalent CDD system $S'$ and
solving $S'$ directly.
%
In general, $S'$ may have size exponential in the size of
$S$.
%
In practice, a suitable equivalent system can be generated using
heuristics analogous to those used to generate compact verification
conditions of hierarchical programs~\cite{flanagan01,lal-qadeer15}.
%
A general procedure for constructing a CDD expansion of a given CHC
system is given in \autoref{app:cons-cdd}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "p"
%%% End: