### 2018-04-14 paper/abstract.tex
worked in cvoss comments

\begin{abstract}
  % problem, context
  Recursion-free Constrained Horn Clauses (CHCs) are logic-programming
  problems that can model safety properties of programs with bounded
  iteration and recursion.
  %
  In addition, many CHC solvers reduce recursive systems to a series
  of recursion-free CHC systems that can each be solved efficiently.

  % contribution:
  In this paper, we define a novel class of recursion-free systems,
  named \emph{Clause-Dependence Disjoint} (CDD), that
  generalizes classes defined in previous work.
  %
  The advantage of this class is that many CDD systems are smaller
  than systems which express the same constraints but are part of a
  different class.
  %
  This advantage in size allows CDD systems to be solved more
  efficiently than their counterparts in other classes.
  %
  We implemented a CHC solver named
  \sys.  \sys~solves arbitrary CHC systems by reducing the input to a
  series of CDD systems.
  %
  Our evaluation indicates that \sys~outperforms state-of-the-art
  implementations in many practical cases.
\end{abstract}