### 2018-04-27 paper/fig/hierarchy2.tex

% \scalebox{0.75}{%
% }

\begin{tikzpicture}
\tikzstyle{every node}=[draw=black, thin, ellipse]
  \begin{scope}[fill opacity=0.1,text opacity=1]

  \draw node[minimum width=6.2cm,minimum height=6cm,
    label={[anchor=north, inner sep=5pt]north:
      \begin{tabular}{c}Recursion-Free\\\textsc{co-NEXPTIME Complete}\end{tabular}} ]
    at (0,0)
    {};

  \draw node[minimum width=5.5cm,minimum height=4.5cm,
    label={[anchor=south, inner sep=0pt]south:
      \begin{tabular}{c}Directly Solvable\\\textsc{co-NP}\end{tabular}} ]
    at (0,-6mm)
    {};

  \draw node[very thick, minimum width=5cm,minimum height=3cm,
    label={[anchor=north, inner sep=5pt]north:CDD}]
    at (0,-1mm)
    {};

  \draw node[minimum width=2.7cm,minimum height=2cm]
    at (10mm,-2mm)
    {\begin{tabular}{c}Body\\Disjoint\end{tabular}};

  \draw node[minimum width=2.7cm,minimum height=2cm]
    at (-10mm,-2mm)
    {Linear};

    % \begin{tabular}{c}Directly Solvable\\\textsc{co-NP}\end{tabular}} ]
  \end{scope}
\end{tikzpicture}